Comments on the revised version

### 2018-03-08 paper/paper.tex
revision2 first draft.

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set. The increasing amount and
diversity of data generated by large scale platforms motivate the community to
study more sophisticated policies that could leverage these data. During the
last few years, there was an explosion in the number of works at the interface
of the HPC and Big Data fields, dealing with learning algorithms. Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
times~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
evaluate the impact of such techniques on the performances of the resource
manager.

@@

We further assume that there is a certain regularity among periods, i.e.\ that
the distribution of the job submission process does not radically differ
between consecutive periods. This assumption is validated in the study
presented in~\cite{jsspp17}. It further entails that the behavior of a policy
during the previous periods reflects its behavior on the current one, so that
the selection of a policy can be based on its past behavior. However, there may
be some variability between different periods for certain cost
metrics~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:variability} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all 7
policies considered. This indicates that when the cost metric is averaged
over different periods, there is a trade-off to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

@@

Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the
period $\Delta_t$ ($0 \le t $), and $P_i$ ($1 \le i \le 12$) one reordering
policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$
during period $\Delta_t$ is defined as:
%
\begin{equation} \label{eq:cost-exact} w(t;P_i) = \sum_{j=1}^{l(\Delta_t)}
\mbox{Wait}_j^{P_i} \end{equation}
%
where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of
the waiting time for job $j$ according to policy $P_i$. The estimation of the
cost of a policy over all the periods preceding the current period can then be
defined as:
%
\begin{equation} \label{eq:tot-cost-exact} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w(t;P_i) \end{equation}
%
where $\lambda \in [0,1]$ is a decaying discount factor that can be used to
privilege recent history (\textit{i.e.} recent periods). The above formulation
is general and covers several possible situations : From stationary job
distributions where $\lambda$ can be set to 1 to non-stationary distributions
where the more recent history has to be privileged. As explained in Section
6.1, in our experiments $\lambda$ is set to 1 as the distributions considered
are stationary but the reader has to bear in mind that our framework can be use
on different distributions by resorting to different values of $\lambda$ that can
be set by cross-validation.

@@

  \caption{Illustration of the online simulation-based strategy. Every week,
  the wokload from the past week is used to run simulations using every
  reordering policy. The "model" $w$ is updated according
  to~\eqref{eq:tot-cost-exact} (or~\eqref{eq:tot-cost-noisy} in the Noisy case
  from Section~\ref{sub:noisy}) and the policy for the next week is chosen
  according to~\eqref{eq:select-exact}}.

@@

  \caption{Illustration of the bandit-based strategy. Every week,
  the wokload from the past week is used to run simulations using every
  reordering policy. The "model" $w$ is updated according
  to~\eqref{eq:tot-cost-greedy} and the policy for the next week is chosen
  according to~\eqref{alg:epsilon}. We denote by $P_{\mbox{bandit},t}$ the
  policy chosen by the bandit at week $t$.
  }.

@@

>>>>

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set. The increasing amount and
diversity of data generated by large scale platforms motivate the community to
study more sophisticated policies that could leverage these data. During the
last few years, there was an explosion in the amount of work at the interface
of the HPC and Big Data fields, dealing with learning algorithms. Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
times~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
evaluate the impact of such techniques on the performances of the resource
manager.

@@

We further assume that there is a certain regularity among periods, i.e.\ that
the distribution of the job submission process does not radically differ
between consecutive weekly periods. This assumption is validated in the study
presented in~\cite{jsspp17}. It further entails that the behavior of a policy
during the previous periods reflects its behavior on the current one, so that
the selection of a policy can be based on its past behavior. However, there may
be some variability between different periods for certain cost
metrics~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:variability} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all 7
policies considered. This indicates that when the cost metric is averaged
over different periods, there is a trade-off to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

@@

Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the
period $\Delta_t$ ($0 \le t $), and $P_i$ ($1 \le i \le 12$) one reordering
policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$
during period $\Delta_t$ is defined as:
%
\begin{equation} \label{eq:cost-exact} w(t;P_i) = \sum_{j=1}^{l(\Delta_t)}
\mbox{Wait}_j^{P_i} \end{equation}
%
where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of
the waiting time for job $j$ according to policy $P_i$. The estimation of the
cost of a policy over all the periods preceding the current period can then be
defined as:
%
\begin{equation} \label{eq:tot-cost-exact} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w(t;P_i) \end{equation}
%
where $\lambda \in [0,1]$ is a decaying discount factor that can be used to
privilege recent history (\textit{i.e.} recent periods). The above formulation
is general and covers several possible situations : From stationary job
distributions where $\lambda$ can be set to 1 to non-stationary distributions
where the more recent history has to be privileged. As explained in
Section~\ref{ssub:resampling}, in our experiments $\lambda$ is set to 1 as the
distributions considered are stationary but the reader has to bear in mind that
our framework can be use on different distributions by resorting to different
values of $\lambda$ that can be set by cross-validation.

@@

  \caption{Illustration of the online simulation-based strategy. Every week,
  the workload from the past week is used to run simulations using every
  reordering policy. The "model" $w$ is updated according
  to~\eqref{eq:tot-cost-exact} (or~\eqref{eq:tot-cost-noisy} in the Noisy case
  from Section~\ref{sub:noisy}) and the policy for the next week is chosen
  according to~\eqref{eq:select-exact}}.

@@

  \caption{Illustration of the bandit-based setting. Every week,
  the wokload from the past week is used to run simulations using every
  reordering policy. The "model" $w$ is updated according
  to~\eqref{eq:tot-cost-greedy} and the policy for the next week is chosen
  according to~\eqref{alg:epsilon}. We denote by $P_{\mbox{bandit},t}$ the
  policy chosen by the bandit at week $t$.
  }.

@@

\textbf{SQF versus online tuning.}
Table~\ref{tab:full} shows that the SQF policy
works quite well on the traces considered here on average, and under this
resampling scheme. The variability however, is high, and this policy is not the
best on all traces. It interesting to have an experimental evaluation of
exactly how much one gains on average by using an online policy versus a
leave-one-trace-out majority vote policy choice.  However, a good estimation of
this quantity is tricky. Indeed, we expect the change in the relative
performance of policies to increase with two factors. First, simply adding
workload traces to our selection will naturally show more cases where behavior
is different. Second, factoring in the topology model of the machines should
increase the heterogeneity between platforms and therefore increase the
dependence of the simulation performance on the trace/platform. Such a study
would be much more tedious. Here the need for adaptivity is actually
understated by the simulation approach.

While we can not conclude on how much one gains by not using SQF in this case,
we can comment on the added complexity and computing overhead of using the
simulation-based and bandit online strategies.  In both these cases, the
scheduling decisions do not have an overhead, as the policy is fixed during the
period (one day, or week). The bandit strategy has a negligible computational
overhead at period end. Its two computational operations are maintaining a
sequential average and generating a pseudo-random number.  The simulation-based
strategy requires to simulate the system K times every period end, where K is
the number of alternative policies considered (here, K=12). Using our
simulator, This takes a few seconds. Using a more precise simulator with
topological modeling, one can argue that the overhead is still manageable since
these simulations can be easily done in parallel. The most important constraint
is that these simulations should not take too much time compared to the period
size, as to not delay the policy switch decision too much.

### 2018-03-08 paper/revision2.tex
revision2 first draft.

\usepackage{listings}
\usepackage{fullpage}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

@@

Let us first sincerely thank the reviewers for their relevant and in-depth
remarks once again. Their comments helped us further the discussion of the
results, and find a typographical error that passed our scrutiny.

@@

@@

We regret that the detailed answer of the previous rebuttal did not convince you. We disagree with this statement. We indeed
provide a new scheduling algorithm, which adapts to the situation by selecting the correct reordering policy. This algorithm
results in significant improvements over existing methods.
We thank you for your in-depth remarks once again. We integrated these comments into the minor revision.

@@

Right, this discussion is warranted. We added paragraph X.

@@

Indeed, the SQF policy works quite well on the traces considered here on
average, and under this resampling scheme. The variability however, is high,
and it is not the best on other traces.  It would indeed be interesting to have
an experimental evaluation of exactly how much one gains on average by choosing
the bandit over a leave-one-trace-out majority vote policy choice.

However, a good estimation of this quantity requires using more traces, and
perhaps factoring in the topology model of the machines, which should increase
the heterogeneity between platforms and therefore increase the dependence of
the simulation performance on the trace/platform. Such a study would be much
more tedious. Here, we prove the feasibility of the approach. The need for
adaptivity is actually understated by our simulation approach, but in all
honesty we don't believe to be able to measure this rigorously.

@@

Right, we need to make this clearer. We added this discussion to paragraph X.

In both these cases, the scheduling decisions do not have an overhead, as the
policy is fixed during the period (one day, or week).

The bandit strategy has a negligible computational overhead at period end. Its two
computational operations are maintaining a sequential average and generating a
pseudo-random number.

The simulation-based strategy requires to simulate the system K times every
period end, where K is the number of alternative policies considered (here,
K=12). Using our simulator, This takes a few seconds. Using a more precise
simulator with topological modeling, one can argue that the overhead is still
manageable since these simulations can be easily done in parallel. The most
important constraint is that these simulations should not take too much time
compared to the period size, as to not delay the policy switch decision too
much.

>>>>

\usepackage{csquotes}
\usepackage{listings}
\usepackage{fullpage}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

@@

Let us first sincerely thank the reviewers for their relevant and in-depth
remarks once again. Their comments helped us further the discussion of the
results, and find a typographical error in the presentation of the results that
passed our scrutiny.

@@

\medskip

@@

We regret that the detailed answer to your comments did not convince you. We disagree with this statement. We indeed
provide a new scheduling algorithm, which adapts to the situation by selecting the correct reordering policy. This algorithm
results in significant improvements over existing methods.
We thank you for your in-depth remarks once again. We integrated these comments into the minor revision.

\medskip

@@

@@

You are right, the discussion you suggest is warranted. We added it to
paragraph \textbf{SQF versus online tuning} in Section 6.3.2:

\begin{displayquote}
Table~3 shows that the SQF policy
works quite well on the traces considered here on average, and under this
resampling scheme. The variability however, is high, and this policy is not the
best on all traces. It interesting to have an experimental evaluation of
exactly how much one gains on average by using an online policy versus a
leave-one-trace-out majority vote policy choice.  However, a good estimation of
this quantity is tricky. Indeed, we expect the change in the relative
performance of policies to increase with two factors. First, simply adding
workload traces to our selection will naturally show more cases where behavior
is different. Second, factoring in the topology model of the machines should
increase the heterogeneity between platforms and therefore increase the
dependence of the simulation performance on the trace/platform. Such a study
would be much more tedious. Here the need for adaptivity is actually
understated by the simulation approach.
\end{displayquote}

\medskip

@@

Right, we need to make this clearer. We added this discussion to paragraph
\textbf{SQF versus online tuning} in Section 6.3.2:

\begin{displayquote}
While we can not conclude on how much one gains by not using SQF in this case,
we can comment on the added complexity and computing overhead of using the
simulation-based and bandit online strategies. In both these cases, the
scheduling decisions do not have an overhead, as the policy is fixed during the
period (one day, or week). The bandit strategy has a negligible computational
overhead at period end. Its two computational operations are maintaining a
sequential average and generating a pseudo-random number.  The simulation-based
strategy requires to simulate the system K times every period end, where K is
the number of alternative policies considered (here, K=12). Using our
simulator, This takes a few seconds. Using a more precise simulator with
topological modeling, one can argue that the overhead is still manageable since
these simulations can be easily done in parallel. The most important constraint
is that these simulations should not take too much time compared to the period
size, as to not delay the policy switch decision too much.
\end{displayquote}

\medskip
